---
title: "Practical Machine Learning"
author: "SaidatulRahah"
date: "Wednesday, December 16, 2015"
output: html_document
---
##Practice Machine learning project

#Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here

#Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


```{r echo=FALSE}
library(caret)
library(kernlab)
library(ggplot2)
library(lattice)
#library(e1071)
#library(gbm)
library(rpart)
library(rpart.plot)
```
```{r warning = FALSE}


# load data
train <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"), row.names = 1)
test <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"), row.names = 1)
table(train$classe)
```
##Preprocessing

#Partitioning the training set

We separate our training data into a training set and a validation set so that we can validate our model.  the training data set is partionned into 2 sets: subTraining (80%) and subTest (20%).
```{r}
set.seed(123456)
trainset <- createDataPartition(train$classe, p = 0.8, list = FALSE)
Training <- train[trainset, ]
Validation <- train[-trainset, ]
```
#Feature selection

First we clean up near zero variance features, columns with missing values and descriptive fields.
```{r}
# exclude near zero variance features
nzvcol <- nearZeroVar(Training)
Training <- Training[, -nzvcol]

# exclude columns with m40% ore more missing values exclude descriptive
# columns like name etc
cntlength <- sapply(Training, function(x) {
    sum(!(is.na(x) | x == ""))
})
nullcol <- names(cntlength[cntlength < 0.6 * length(Training$classe)])
descriptcol <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")
excludecols <- c(descriptcol, nullcol)
Training <- Training[, !names(Training) %in% excludecols]
```

The variable "classe" contains 5 levels: A, B, C, D and E. A plot of the outcome variable will allow us to see the frequency of each levels in the subTraining data set and compare one another.

```{r}
plot(Training$classe, col="blue", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")

```

#Model Train 1: Random Forest

We will use random forest as our model as implemented in the randomForest package by Breiman's random forest algorithm (based on Breiman and Cutler's original Fortran code) for classification and regression.

```{r}
library(randomForest)
rfModel <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 10)
ptraining <- predict(rfModel, Training)
print(confusionMatrix(ptraining, Training$classe))
```
Obviously our model performs excellent against the training set, but we need to cross validate.

#Model Train 2: Decision Tree
```{r}
model1 <- rpart(classe ~ ., data=Training, method="class")

# Predicting:
prediction1 <- predict(model1, Validation, type = "class")

# Plot of the Decision Tree
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)

# Test results on our Validation data set:
confusionMatrix(prediction1, Validation$classe)
```


#Validation set accuracy (Out-of-Sample)

Let us now see how our model performs on the cross validation set that we held out from training.
```{r}
pvalidation <- predict(rfModel, Validation)
print(confusionMatrix(pvalidation, Validation$classe))
```
The cross validation accuracy is 99.5% and the out-of-sample error is therefore 0.5% so our model performs rather good.


#Test set prediction

The prediction of our algorithm for the test set is:
```{r}
ptest <- predict(rfModel, test)
ptest

```


#Conclusion

As a conclusion the Random Forest method worked very well as compared to Decision Tree. Accuracy for Random Forest model was 0.995 (95% CI: (0.993, 0.997)) compared to 0.749 (95% CI: (0.736, 0.763)) for Decision Tree model. The random Forest model is choosen. The Confusion Matrix achieved 99.5% accuracy. This model will be used for the final calculations.


Prepare the submission.
```{r}
answers <- as.vector(ptest)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)
```


